Estimating duration in the presence of misclassification

### 2024-03-19 MCMC-appendix.tex
spelling / grammar check

\section{Metropolis-Hastings}

A common way of constructing a Markov chain with $p(\psi \mid y)$ as its stationary distribution is using the Metropolis-Hastings algorithm~\autocite{hastingsMCMC}.
The algorithm requires a proposal distribution $q(\psi' \mid \psi)$, which proposes that the chain moves to a new state $\psi'$ from its current state $\psi$; $\pi(\psi)$, a calculation of $p(\psi \mid y)$ up to some constant.
At each iteration, a proposal is draw from $q$ and then accepted or rejected based on the \emph{acceptance ratio}, computed using $\pi$.
The full algorithm is given in \cref{inc-prev:algorithm:MH}, with further details in available in a wide-range of sources~\autocites[e.g.][]{brooksMCMCNotes}[chapter 11]{gelmanBDA}.
\begin{algorithm}
 set $\psi^0$ to some initial value\;
 \For{$i = 1, \dots, M$}{
  sample a proposal $\psi' \sim q(\psi' \mid \psi)$ \;
  calculate the acceptance probability $\alpha(\psi, \psi') = \min \left( 1, \frac{\pi(\psi') q(\psi \mid \psi')}{\pi(\psi) q(\psi' \mid \psi)} \right)$ \;
  with probability $\alpha(\psi, \psi')$ set $\psi^i$ to $\psi'$, otherwise set $\psi^i$ to $\psi^{i-1}$ \;
 }
 \caption{The Metropolis-Hastings algorithm.}
 \label{inc-prev:algorithm:MH}
\end{algorithm}

The proposal distribution $q$ is most commonly a (possibly multivariate) Normal distribution centred on $\psi$; this scheme is used in \cref{E-SEIR}.
Tuning the proposal distribution to propose ``good'' moves is a key part of MCMC.
The ideal proposal distribution is the posterior of interest, however, this is impossible because that is what MCMC is trying to estimate~\autocite[296]{gelmanBDA}.
One alternative is \emph{adaptive} algorithms which tune the proposal distribution based on previous draws in an attempt to approximate the posterior; I use such a scheme in \cref{E-SEIR:sec:inference}.

@@

MCMC using a Normal proposal distribution can be inefficient.
In particular, it often poorly explores the tails of the posterior in finite time~\autocite[e.g.][]{turitsynIrreversible}.
\emph{Hamiltonian Monte Carlo} (HMC)~\autocite{nealMCMC,nealImproved,duaneHybrid} improves the proposals by using information from the gradient of the posterior density (taken with respect to the parameters).
HMC simulates a physical system travelling through the posterior space with momentum.
The momentum increases in the direction the posterior density increases, using the gradient.
This means that the sampler tends to propose moves with higher posterior density.
The Metropolis-Hastings algorithm is then used to accept or reject the proposed move.

>>>>

\section{Metropolis--Hastings}

A common way of constructing a Markov chain with $p(\psi \mid y)$ as its stationary distribution is using the Metropolis--Hastings algorithm~\autocite{hastingsMCMC}.
The algorithm requires a proposal distribution $q(\psi' \mid \psi)$, which proposes that the chain moves to a new state $\psi'$ from its current state $\psi$; $\pi(\psi)$, a calculation of $p(\psi \mid y)$ up to some constant.
At each iteration, a proposal is draw from $q$ and then accepted or rejected based on the \emph{acceptance ratio}, computed using $\pi$.
The full algorithm is given in \cref{inc-prev:algorithm:MH}, with further details in available in a wide-range of sources~\autocites[e.g.][]{brooksMCMCNotes}[chapter 11]{gelmanBDA}.
\begin{algorithm}
 set $\psi^0$ to some initial value\;
 \For{$i = 1, \dots, M$}{
  sample a proposal $\psi' \sim q(\psi' \mid \psi)$ \;
  calculate the acceptance probability $\alpha(\psi, \psi') = \min \left( 1, \frac{\pi(\psi') q(\psi \mid \psi')}{\pi(\psi) q(\psi' \mid \psi)} \right)$ \;
  with probability $\alpha(\psi, \psi')$ set $\psi^i$ to $\psi'$, otherwise set $\psi^i$ to $\psi^{i-1}$ \;
 }
 \caption{The Metropolis--Hastings algorithm.}
 \label{inc-prev:algorithm:MH}
\end{algorithm}

The proposal distribution $q$ is most commonly a (possibly multivariate) normal distribution centred on $\psi$; this scheme is used in \cref{E-SEIR}.
Tuning the proposal distribution to propose ``good'' moves is a key part of MCMC.
The ideal proposal distribution is the posterior of interest, however, this is impossible because that is what MCMC is trying to estimate~\autocite[296]{gelmanBDA}.
One alternative is \emph{adaptive} algorithms which tune the proposal distribution based on previous draws in an attempt to approximate the posterior; I use such a scheme in \cref{E-SEIR:sec:inference}.

@@

MCMC using a normal proposal distribution can be inefficient.
In particular, it often poorly explores the tails of the posterior in finite time~\autocite[e.g.][]{turitsynIrreversible}.
\emph{Hamiltonian Monte Carlo} (HMC)~\autocite{nealMCMC,nealImproved,duaneHybrid} improves the proposals by using information from the gradient of the posterior density (taken with respect to the parameters).
HMC simulates a physical system travelling through the posterior space with momentum.
The momentum increases in the direction the posterior density increases, using the gradient.
This means that the sampler tends to propose moves with higher posterior density.
The Metropolis--Hastings algorithm is then used to accept or reject the proposed move.

### 2024-03-19 cis-imperfect-testing.tex
spelling / grammar check

In \cref{E-perf-test}, a framework for analysing the CIS to produce estimates of the duration of positivity was developed.
However, initial application of this framework produced an implausibly short estimate.
In this chapter I identify this issue to be the result of false negatives, a form of misclassification bias, within the CIS (\cref{imperf-test:sec:problem}).
I then propose several models of false negatives (\cref{imperf-test:sec:simulate}) and extend the model of \cref{E-perf-test} to incorporate one of them (\cref{imperf-test:sec:modelling})
After showing through simulation that the new model recovers the true duration distribution (\cref{imperf-test:sec:sim-study-results}), I apply it to the CIS data (\cref{imperf-test:sec:application}).
Finally, I discuss the results and further work in \cref{imperf-test:sec:discussion}.

@@

Comparing repeated data simulations, using the setup of \cref{E-perf-test:sec:simulation-study}, to the observed CIS data, suggests that this is due to the high frequency of single positive episodes in the CIS data (compare the histogram with $\psens=1$ and the vertical line in \cref{imperf-test:fig:sim-single-pos}).
A single positive episode is an episode containing exactly one positive test; intuitively, these cause estimates to be shorter because the lower bound on the length of time the episode is 1 day, the shortest possible.
Specifically, if $j$ is a single positive episode then $j$'s admissible region (as defined in \cref{E-perf-test:sec:model}) includes $b_j = e_j = r_j^{(b)} = l_j^{(e)}$.
In this case, the beginning and end day of the episode are both equal to the day of this one positive test.
If this is true, then the duration is $d_j = e_j - b_j + 1 = 1$.
Additionally, short episodes are very likely to be missed, which means that the design of the CIS cannot rule out many of them occurring.
% This produced the same pattern, that is estimating too many short episodes (not shown).

\Textcite{shenNonparametrica}, generalising \textcite{panNote}, studied the situation when only the terminating event is interval censored (\ie singly censored data) and left truncated.
They showed that the non-parameteric maximum likelihood estimator is inconsistent if single positive episodes occur, and the survival function is severely biased downwards.

@@

To assess the sensitivity of the procedure to a time-varying test sensitivity, it will be useful to simulate from a model with a varying test sensitivity as a function of days since the start of the infection episode.
Through consideration of \cref{imperf-test:fig:bounding-cis-sensitivity}, and the interaction of false negatives due to viral load in \cref{E-ATACCC}'s results, I propose the following model.
\begin{equation}
  p_\text{sens}(t) = \begin{cases}
    0.9 - \frac{0.9-0.5}{50}t &t \leq 50 \\
    0.5 &t > 50
  \end{cases}
  \label{imperf-test:eq:variable-test-sensitivity}
\end{equation}
where $t$ is the number of days since infection.
Therefore, this model consists of a linear decline in test sensitivity, before a constant minimum point.

@@

I will modify $p_{ia}$ to allow the negative test following the last positive to be a false negative.
If it is a false negative, then I will consider the episode's end right-censored.
However, if it is a true negative, then the episode's end is interval-censored, as previously.
A mixture of these scenarios forms the episode's likelihood contribution, with the mixture probability determined by the test sensitivity.

@@

For tractability, assume that the negative test bounding the start of the episode, on day $l_{j(i)}^{(b)}-1$, is a true negative.
This assumption is reasonable because, since a positive test follows at $r_{j(i)}^{(b)}$, the negative at $l_{j(i)}^{(b)}-1$ is likely early in the infection and the test sensitivity is high early in an infection.
Therefore, this is unlikely to be a false negative because, if the individual was detectable at this time, they are in a period with high test sensitivity.
This true negative, occurs with probability 1, and hence does not contribute to the likelihood.
Denote the remaining tests considered as $\sched'_i = \{ t \in \sched_i \ssep r_{j(i)}^{(b)} \leq t \leq r_{j(i)}^{(e)} + 1 \}$, and their results by $\vec{y_{j(i)}} = \{ y_{j(i)}(t) \ssep t \in \sched'_i \}$ where $y_{j(i)}(t) = 1$ if the test on day $t$ is positive and 0 otherwise.

@@

Consider the negative test at $r_{j(i)}^{(e)}+1$, the first negative after the start of the episode which may be a false negative.
It is a false negative if and only if the episode ends after the test, \ie $e_{j(i)} > r_{j(i)}^{(e)}$.
I proceed by considering whether this is the case, and considering the case with $b_{j(i)}$ known; $b_{j(i)}$ will later be integrated out.

First, if $e_{j(i)} \leq r_{j(i)}^{(e)}$, meaning that the test at $r_{j(i)}^{(e)}+1$ is a true negative and the end of the episode is interval-censored as in the previous chapter.
% In this case, the test at $r_{j(i)}^{(e)} + 1$ is a true negative, as are all other tests not in $\sched'_i$.
The true negative occurs with probability 1, by the assumption of no false positives.
\begin{align}
&p(\vec{y_{j(i)}}, e_{j(i)} \leq r_{j(i)}^{(e)} | b_{j(i)}, p_\text{sens}, \theta) \\
&= p(\vec{y_{j(i)}}, l_{j(i)}^{(e)} \leq e_{j(i)} \leq r_{j(i)}^{(e)} | t_i, b_{j(i)}, p_\text{sens}, \theta) \\ % &\text{as no false positives}
&= p(\vec{y_{j(i)}} \mid l_{j(i)}^{(e)} \leq e_{j(i)} \leq r_{j(i)}^{(e)}, t_i, b_{j(i)}, p_\text{sens}, \theta) p(l_{j(i)}^{(e)} \leq e_{j(i)} \leq r_{j(i)}^{(e)} | t_i, b_{j(i)}, p_\text{sens}, \theta) \\
&= \left( \prod_{t \in \sched'_i} p_\text{sens}^{y_{j(i)}(t)} (1 - p_\text{sens})^{(1 - y_{j(i)}(t))} \right) \left( S_\theta(l_{j(i)}^{(e)} - b_{j(i)} - 1) - S_\theta(r_{j(i)}^{(e)} - b_{j(i)} - 1) \right)
\label{imperf-test:eq:ll-ei-lt-ri}
\end{align}

Second, if $e_{j(i)} > r_{j(i)}^{(e)}$.
In this case, the test at $r_{j(i)}^{(e)}$ is a false negative, occurring with probability $(1 - p_\text{sens})$.
To avoid having to consider tests after $r_{j(i)}^{(e)}$, which could greatly complicate the likelihood, I model this case as the episode being right-censored at $r_{j(i)}^{(e)}$.
Taking the same approach as before:
\begin{align}
&p(\vec{y_{j(i)}}, e_{j(i)} > r_{j(i)}^{(e)} | b_{j(i)}, p_\text{sens}, \theta) \\
&= p(\vec{y_{j(i)}} \mid e_{j(i)} > r_{j(i)}^{(e)}, t_i, b_{j(i)}, p_\text{sens}, \theta) p(e_{j(i)} > r_{j(i)}^{(e)} | t_i, b_{j(i)}, p_\text{sens}, \theta) \\
&= \left( \prod_{t \in \sched'_i} p_\text{sens}^{y_{j(i)}(t)} (1 - p_\text{sens})^{(1 - y_{j(i)}(t))} \right) (1 - p_\text{sens}) S_\theta(r_{j(i)}^{(e)} - b_{j(i)} - 1)
\label{imperf-test:eq:ll-ei-gt-ri}
\end{align}

@@

For each simulated dataset, I estimate the survival function using the likelihood proposed in \cref{imperf-test:sec:modelling} and assuming $\psens$ equals 0.6, 0.8, or 1.0.
If the dataset was simulated using a constant $\psens$, and that value of $\psens$ matches the value assumed during inference, I refer to $\psens$ as being correctly specified; otherwise I refer to it as misspecified.
However, there is still some misspecification of the model to false negatives owing to the simplifying assumptions made.
The amount that the simplifying assumptions are violated increases as $\psens$ decreases.

@@

When $\psens = 0.8$ and is correctly specified, the model recovers the true survival time well (see \cref{imperf-test:fig:constant-test-sensitivity}(B)).
The informative prior, in comparison to the vague prior, helps to overcome the misspecification due to the simplifying assumptions, moving the estimated survival function closer to its true survival time.
However, when $\psens = 0.6$, this is no longer the case (see \cref{imperf-test:fig:constant-test-sensitivity}(A)).
This is likely caused by too large a violation of the simplifying assumptions made in \cref{imperf-test:sec:modelling}.
\begin{figure}
  % \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{cis-imperfect-testing/sim-constant-sensitivity}}
  \includegraphics[width=\textwidth]{cis-imperfect-testing/sim-constant-sensitivity}
  \caption[Simulation study results with constant test sensitivity]{%
    Posterior (median and 95\% credible interval) of the survival time for the simulation study with a correctly specified test sensitivity.
    True survival time shown in black.
  }
  \label{imperf-test:fig:constant-test-sensitivity}
\end{figure}

Next, I considered the consequence of $\psens$ being misspecified, using the vague prior; the better performing prior in the correctly specified case.
If the test sensitivity is misspecified, that is the assumed value for $p_\text{sens}$ in \cref{imperf-test:eq:pit-prime,imperf-test:eq:pit-prime} is different to the value used in the simulation, then the survival time will be biased (see \cref{imperf-test:fig:misspecified-test-sensitivity}).
If $\psens$ is misspecified and too low (see \cref{imperf-test:fig:misspecified-test-sensitivity}(A)), then the posterior estimate initially follows the true value but then separates.
The number of episodes inferred to have truly ended by the first negative is too high, and hence the survival time is overestimated.
This effect dominates over the opposing bias of overestimating the number of missed episodes.
The opposite occurs if the test sensitivity is assumed to be too high, although the posterior moves away from the truth earlier (see \cref{imperf-test:fig:misspecified-test-sensitivity}(C)).
\begin{figure}
  % \makebox[\textwidth][c]{
    \includegraphics[width=\textwidth]{cis-imperfect-testing/sim-misspecified-sensitivity}
  \caption[Simulation study results with misspecified test sensitivity]{%
    Posterior (median and 95\% credible interval) of the survival time for the simulation study with a possibly misspecified test sensitivity.
    True survival time shown in black.
    All simulations use a constant test sensitivity of 0.8, but the inference procedure assumes different values, as per the key.
    Hence, (B) is correctly specified but (A) and (C) have misspecified $\psens$.
  }
  \label{imperf-test:fig:misspecified-test-sensitivity}
\end{figure}

@@

The qualitative increase in long episodes is robust to the choice of prior for $\ntot$, the assumed value for $\psens$ (see \cref{imperf-test:fig:cis-sensitivity}), and the choice of prior for the hazards, $\lambda_t$ (see \cref{E-perf-test:sec:parameters-priors}).
However, the survival proportion over the first 4 weeks is sensitive to these choices.
The estimate using a test sensitivity of 0.8 and $\NBc(\mu\inform, r\inform)$ give a median survival time most similar to that of \cref{E-ATACCC}.
\begin{figure}
  \centering \includegraphics{cis-imperfect-testing/CIS_final}
  \caption{Duration estimates using CIS and ATACCC data}
  \label{imperf-test:fig:cis-estimates}
\end{figure}
\begin{figure}
  \vspace{-2.5cm}
  \makebox[\textwidth][c]{\includegraphics{cis-imperfect-testing/CIS_vary}}
  \caption{%
    (A-C) Sensitivity of CIS estimates to the value of $r$, the dispersion of the prior for $\ntot$ (a larger $r$ means a more informative prior) when $\psens = 0.8$.
    (D-F) Sensitivity of CIS estimates to the value of $\psens$ when $r = r\inform$.
    A and D: median survival time, in comparison to the ATACCC-based estimate of \cref{E-ATACCC} (shown in red).
    B and E: survival at day 50, $S_\theta(50)$.
    C and F: full survival curves out to day 100.
  }
  \label{imperf-test:fig:cis-sensitivity}
  \todo[inline]{Optional improvement: Could highlight the "preferred" model which exists in both rows and is the one used subsequently, and/or include ATACCC in the far-right hand column}
\end{figure}

The the estimates are most sensitive to the choice of $r$, the strength of the prior on $\ntot$.
A very weak, almost uninformative, prior on this quantity causes the posterior estimate to be much higher than the estimate from \textcite{birrellRTM2}.
When increasing the prior's strength, the posterior estimate moves towards the prior smoothly, as expected (see \cref{imperf-test:fig:ntot}).
As discussed previously, the \cref{E-ATACCC} estimates are reliable for the first 2--3 weeks, notably including the median time.
The median using $r\inform$ matched \cref{E-ATACCC}'s median estimate and is a principled choice because it is based directly on the prior work \textcite{birrellRTM2}.
Therefore, I use this estimate in the application to the CIS data.
\begin{figure}
  \centering \includegraphics{cis-imperfect-testing/CIS_ntot}
  \caption{How the posterior estimate of $\ntot$ changes with the value of $r$ in the prior on $\ntot$.}
  \label{imperf-test:fig:ntot}
\end{figure}

@@

Estimating the test sensitivity excluding intermittent negatives is not possible because \cref{imperf-test:eq:pia-prime-constant,imperf-test:eq:pit-prime} are both monotonically decreasing in $\psens$; therefore, the likelihood always favours $\psens = 0$ (\ie no true positives).
This aligns with the situation with singly interval-censored, untruncated data, when stopping at the first observed time of the terminating event (which may be misclassified) means that the test sensitivity cannot be estimated~\autocite[e.g.]{titmanMisclassify}.

Estimating a time-varying $\psens$ with $S_\theta$ jointly may cause identifiability issues.
Previous studies have avoided issues by including external information (such as a prior giving the magnitude) on $\psens$, or test results from later follow-up~\autocite[and references therein]{piresIntervalMisclassify}.
However, these studies use a constant $\psens$.
Whether these methods are sufficient for the model to be identifiable with a time-varying $\psens$ needs further investigation.
A simple parameteric form of the test sensitivity, such as that proposed by \textcite{brownBayesian}, may be sufficient to allow identifiability.
In any case, the likelihood, especially $p_{iu}'$, would be substantially complicated by such an addition which may greatly increase the computational cost of inference.

>>>>

In \cref{E-perf-test}, a framework for analysing the CIS to produce estimates of the duration of positivity was developed.
However, the initial application of this framework produced an implausibly short estimate.
In this chapter I identify this issue to be the result of false negatives, a form of misclassification bias, within the CIS (\cref{imperf-test:sec:problem}).
I then propose several models of false negatives (\cref{imperf-test:sec:simulate}) and extend the model of \cref{E-perf-test} to incorporate one of them (\cref{imperf-test:sec:modelling})
After showing through simulation that the new model recovers the true duration distribution (\cref{imperf-test:sec:sim-study-results}), I apply it to the CIS data (\cref{imperf-test:sec:application}).
Finally, I discuss the results and further work in \cref{imperf-test:sec:discussion}.

@@

Comparing repeated data simulations, using the set-up of \cref{E-perf-test:sec:simulation-study}, to the observed CIS data, suggests that this is due to the high frequency of single positive episodes in the CIS data (compare the histogram with $\psens=1$ and the vertical line in \cref{imperf-test:fig:sim-single-pos}).
A single positive episode is an episode containing exactly one positive test; intuitively, these cause estimates to be shorter because the lower bound on the length of time the episode is 1 day, the shortest possible.
Specifically, if $j$ is a single positive episode then $j$'s admissible region (as defined in \cref{E-perf-test:sec:model}) includes $b_j = e_j = r_j^{(b)} = l_j^{(e)}$.
In this case, the beginning and end day of the episode are both equal to the day of this one positive test.
If this is true, then the duration is $d_j = e_j - b_j + 1 = 1$.
Additionally, short episodes are very likely to be missed, which means that the design of the CIS cannot rule out many of them occurring.
% This produced the same pattern, that is estimating too many short episodes (not shown).

\Textcite{shenNonparametrica}, generalising \textcite{panNote}, studied the situation when only the terminating event is interval censored (\ie singly censored data) and left truncated.
They showed that the non-parametric maximum likelihood estimator is inconsistent if single positive episodes occur, and the survival function is severely biased downwards.

@@

To assess the sensitivity of the procedure to a time-varying test sensitivity, it will be useful to simulate from a model with a varying test sensitivity as a function of days since the start of the infection episode.
Through consideration of \cref{imperf-test:fig:bounding-cis-sensitivity}, and the interaction of false negatives due to viral load in \cref{E-ATACCC}'s results, I propose the following model.
\begin{equation}
  p_\text{sens}(t) = \begin{cases}
    0.9 - \frac{0.9-0.5}{50}t &t \leq 50 \\
    0.5 &t > 50
  \end{cases}
  \label{imperf-test:eq:variable-test-sensitivity}
\end{equation}
where $t$ is the number of days since infection.
Therefore, this model consists of a linear decline in test sensitivity before a constant minimum point.

@@

I will modify $p_{ia}$ to allow the negative test following the last positive to be a false negative.
If it is a false negative, then I will consider the episode's end right censored.
However, if it is a true negative, then the episode's end is interval censored, as previously.
A mixture of these scenarios forms the episode's likelihood contribution, with the mixture probability determined by the test sensitivity.

@@

For tractability, assume that the negative test bounding the start of the episode, on day $l_{j(i)}^{(b)}-1$, is a true negative.
This assumption is reasonable because, since a positive test follows at $r_{j(i)}^{(b)}$, the negative at $l_{j(i)}^{(b)}-1$ is likely early in the infection and the test sensitivity is high early in an infection.
Therefore, this is unlikely to be a false negative because, if the individual was detectable at this time, they are in a period with high test sensitivity.
This true negative occurs with probability 1, and hence does not contribute to the likelihood.
Denote the remaining tests considered as $\sched'_i = \{ t \in \sched_i \ssep r_{j(i)}^{(b)} \leq t \leq r_{j(i)}^{(e)} + 1 \}$, and their results by $\vec{y_{j(i)}} = \{ y_{j(i)}(t) \ssep t \in \sched'_i \}$ where $y_{j(i)}(t) = 1$ if the test on day $t$ is positive and 0 otherwise.

@@

Consider the negative test at $r_{j(i)}^{(e)}+1$, the first negative after the start of the episode which may be a false negative.
It is a false negative if and only if the episode ends after the test, \ie $e_{j(i)} > r_{j(i)}^{(e)}$.
I proceed by considering whether this is the case and considering the case with $b_{j(i)}$ known; $b_{j(i)}$ will later be integrated out.

First, if $e_{j(i)} \leq r_{j(i)}^{(e)}$, meaning that the test at $r_{j(i)}^{(e)}+1$ is a true negative and the end of the episode is interval censored as in the previous chapter.
% In this case, the test at $r_{j(i)}^{(e)} + 1$ is a true negative, as are all other tests not in $\sched'_i$.
The true negative occurs with probability 1, by the assumption of no false positives.
\begin{align}
&p(\vec{y_{j(i)}}, e_{j(i)} \leq r_{j(i)}^{(e)} | b_{j(i)}, p_\text{sens}, \theta) \\
&= p(\vec{y_{j(i)}}, l_{j(i)}^{(e)} \leq e_{j(i)} \leq r_{j(i)}^{(e)} | t_i, b_{j(i)}, p_\text{sens}, \theta) \\ % &\text{as no false positives}
&= p(\vec{y_{j(i)}} \mid l_{j(i)}^{(e)} \leq e_{j(i)} \leq r_{j(i)}^{(e)}, t_i, b_{j(i)}, p_\text{sens}, \theta) p(l_{j(i)}^{(e)} \leq e_{j(i)} \leq r_{j(i)}^{(e)} | t_i, b_{j(i)}, p_\text{sens}, \theta) \\
&= \left( \prod_{t \in \sched'_i} p_\text{sens}^{y_{j(i)}(t)} (1 - p_\text{sens})^{(1 - y_{j(i)}(t))} \right) \left( S_\theta(l_{j(i)}^{(e)} - b_{j(i)} - 1) - S_\theta(r_{j(i)}^{(e)} - b_{j(i)} - 1) \right)
\label{imperf-test:eq:ll-ei-lt-ri}
\end{align}

Second, if $e_{j(i)} > r_{j(i)}^{(e)}$.
In this case, the test at $r_{j(i)}^{(e)}$ is a false negative, occurring with probability $(1 - p_\text{sens})$.
To avoid having to consider tests after $r_{j(i)}^{(e)}$, which could greatly complicate the likelihood, I model this case as the episode being right censored at $r_{j(i)}^{(e)}$.
Taking the same approach as before:
\begin{align}
&p(\vec{y_{j(i)}}, e_{j(i)} > r_{j(i)}^{(e)} | b_{j(i)}, p_\text{sens}, \theta) \\
&= p(\vec{y_{j(i)}} \mid e_{j(i)} > r_{j(i)}^{(e)}, t_i, b_{j(i)}, p_\text{sens}, \theta) p(e_{j(i)} > r_{j(i)}^{(e)} | t_i, b_{j(i)}, p_\text{sens}, \theta) \\
&= \left( \prod_{t \in \sched'_i} p_\text{sens}^{y_{j(i)}(t)} (1 - p_\text{sens})^{(1 - y_{j(i)}(t))} \right) (1 - p_\text{sens}) S_\theta(r_{j(i)}^{(e)} - b_{j(i)} - 1)
\label{imperf-test:eq:ll-ei-gt-ri}
\end{align}

@@

For each simulated dataset, I estimate the survival function using the likelihood proposed in \cref{imperf-test:sec:modelling} and assuming $\psens$ equals 0.6, 0.8, or 1.0.
If the dataset was simulated using a constant $\psens$, and that value of $\psens$ matches the value assumed during inference, I refer to $\psens$ as being correctly specified; otherwise, I refer to it as misspecified.
However, there is still some misspecification of the model to false negatives owing to the simplifying assumptions made.
The amount that the simplifying assumptions are violated increases as $\psens$ decreases.

@@

When $\psens = 0.8$ and is correctly specified, the model recovers the true survival time well (see \cref{imperf-test:fig:constant-test-sensitivity}(B)).
The informative prior, in comparison to the vague prior, helps to overcome the misspecification due to the simplifying assumptions, moving the estimated survival function closer to its true survival time.
However, when $\psens = 0.6$, this is no longer the case (see \cref{imperf-test:fig:constant-test-sensitivity}(A)).
This is likely caused by too large a violation of the simplifying assumptions made in \cref{imperf-test:sec:modelling}.
\begin{figure}
  % \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{cis-imperfect-testing/sim-constant-sensitivity}}
  \includegraphics[width=\textwidth]{cis-imperfect-testing/sim-constant-sensitivity}
  \caption[Simulation study results with constant test sensitivity]{%
    Posterior (median and 95\% credible interval) of the survival time for the simulation study with a correctly specified test sensitivity.
    The true survival time is shown in black
  }
  \label{imperf-test:fig:constant-test-sensitivity}
\end{figure}

Next, I considered the consequence of $\psens$ being misspecified, I use the model combination prior because it is the better performing prior in the correctly specified case.
If the test sensitivity is misspecified, that is the assumed value for $p_\text{sens}$ in \cref{imperf-test:eq:pit-prime,imperf-test:eq:pit-prime} is different to the value used in the simulation, then the survival time will be biased (see \cref{imperf-test:fig:misspecified-test-sensitivity}).
If $\psens$ is misspecified and too low (see \cref{imperf-test:fig:misspecified-test-sensitivity}(A)), then the posterior estimate initially follows the true value but then separates.
The number of episodes inferred to have truly ended by the first negative is too high, and hence the survival time is overestimated.
This effect dominates over the opposing bias of overestimating the number of missed episodes.
The opposite occurs if the test sensitivity is assumed to be too high, although the posterior moves away from the truth earlier (see \cref{imperf-test:fig:misspecified-test-sensitivity}(C)).
\begin{figure}
  % \makebox[\textwidth][c]{
    \includegraphics[width=\textwidth]{cis-imperfect-testing/sim-misspecified-sensitivity}
  \caption[Simulation study results with misspecified test sensitivity]{%
    Posterior (median and 95\% credible interval) of the survival time for the simulation study with a possibly misspecified test sensitivity.
    The true survival time is shown in black
    All simulations use a constant test sensitivity of 0.8, but the inference procedure assumes different values, as per the key.
    Hence, (B) is correctly specified but (A) and (C) have misspecified $\psens$.
  }
  \label{imperf-test:fig:misspecified-test-sensitivity}
\end{figure}

@@

The qualitative increase in long episodes is robust to the choice of prior for $\ntot$, the assumed value for $\psens$ (see \cref{imperf-test:fig:cis-sensitivity}), and the choice of prior for the hazards, $\lambda_t$ (see \cref{E-perf-test:sec:parameters-priors}).
However, the survival proportion over the first 4 weeks is sensitive to these choices.
The estimate using a test sensitivity of 0.8 and $\NBc(\mu\inform, r\inform)$ give a median survival time most similar to that of \cref{E-ATACCC}.
\begin{figure}
  \centering \includegraphics{cis-imperfect-testing/CIS_final}
  \caption{Duration estimates using CIS and ATACCC data.}
  \label{imperf-test:fig:cis-estimates}
\end{figure}
\begin{figure}
  \vspace{-2.5cm}
  \makebox[\textwidth][c]{\includegraphics{cis-imperfect-testing/CIS_vary}}
  \caption{%
    (A-C) Sensitivity of CIS estimates to the value of $r$, the dispersion of the prior for $\ntot$ (a larger $r$ means a more informative prior) when $\psens = 0.8$.
    (D-F) Sensitivity of CIS estimates to the value of $\psens$ when $r = r\inform$.
    A and D: median survival time, in comparison to the ATACCC-based estimate of \cref{E-ATACCC} (shown in red).
    B and E: survival at day 50, $S_\theta(50)$.
    C and F: full survival curves out to day 100.
  }
  \label{imperf-test:fig:cis-sensitivity}
  \todo[inline]{Optional improvement: Could highlight the "preferred" model which exists in both rows and is the one used subsequently, and/or include ATACCC in the far-right hand column}
\end{figure}

The estimates are most sensitive to the choice of $r$, the strength of the prior on $\ntot$.
A very weak, almost uninformative, prior on this quantity causes the posterior estimate to be much higher than the estimate from \textcite{birrellRTM2}.
When increasing the prior's strength, the posterior estimate moves towards the prior smoothly, as expected (see \cref{imperf-test:fig:ntot}).
As discussed previously, the \cref{E-ATACCC} estimates are reliable for the first 2--3 weeks, notably including the median time.
The median using $r\inform$ matched \cref{E-ATACCC}'s median estimate and is a principled choice because it is based directly on the prior work \textcite{birrellRTM2}.
Therefore, I use this estimate in the application to the CIS data.
\begin{figure}
  \centering \includegraphics{cis-imperfect-testing/CIS_ntot}
  \caption{How the posterior estimate of $\ntot$ changes with the value of $r$ in the prior on $\ntot$.}
  \label{imperf-test:fig:ntot}
\end{figure}

@@

Estimating the test sensitivity excluding intermittent negatives is not possible because \cref{imperf-test:eq:pia-prime-constant,imperf-test:eq:pit-prime} are both monotonically decreasing in $\psens$; therefore, the likelihood always favours $\psens = 0$ (\ie no true positives).
This aligns with the situation with singly interval censored, untruncated data, when stopping at the first observed time of the terminating event (which may be misclassified) means that the test sensitivity cannot be estimated~\autocite[e.g.]{titmanMisclassify}.

Estimating a time-varying $\psens$ with $S_\theta$ jointly may cause identifiability issues.
Previous studies have avoided issues by including external information (such as a prior giving the magnitude) on $\psens$, or test results from later follow-up~\autocite[and references therein]{piresIntervalMisclassify}.
However, these studies use a constant $\psens$.
Whether these methods are sufficient for the model to be identifiable with a time-varying $\psens$ needs further investigation.
A simple parametric form of the test sensitivity, such as that proposed by \textcite{brownBayesian}, may be sufficient to allow identifiability.
In any case, the likelihood, especially $p_{iu}'$, would be substantially complicated by such an addition which may greatly increase the computational cost of inference.
